# Results

## Survey Results

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries from the survey in the dataset after cleaning was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Regarding socio-economic factors, respondents who reported living in a household with an income under CAD 15,000 or above 90,000 represented approximately 50\% of the total number of entries, with 71\% and 75\% of the respondents in each income bracket indicating to have received the first dose of the vaccine, respectively. The age group between 16-24 years had the highest representation in the survey responses (837, 23.6\% of all responses), followed by the age group of those between 65 and over (692, 19.5\% of the total); 73\% of those in the 16 to 24 years age group reported to have received the vaccine, whereas 68\% of those above 65 years indicated the same. The Health Region with the highest number of participants in the survey was Toronto, accounting for 1,342 entries (37.7\%) and a 72\% vaccination rate. Regarding race, individuals that identified as White/Caucasian represented 1313 (37\%) of all entries and had the highest vaccination uptake with 82\% of them indicating to have received the COVID-19 vaccine. On the other hand, those that identified as Indigenous had the lowest vaccine uptake with 60\% indicating to have received the vaccine.


```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey"
#| message: false
#| echo: false

source(here::here("code","data_preparation.R"))

data_summary %>%
  modify_header(label = "**Variable**") %>%
  as_kable_extra(booktabs = TRUE,linesep="") %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down"` argument here.
  kableExtra::kable_styling(latex_options = "repeat_header",font_size = 9)

```

## Multivariate Regression

@tbl-model shows the results of the logistic regression on vaccination status using socio-economic, geographical factors (by Health Region) and their interactions as predictors. The reference groups in each case were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income). Statistically significant odds ratios were identified for individuals with a household income under CAD 25,000 (OR=0.37) and between CAD 25,000-59,999 (OR=0.59), persons who identified as Arab/Middle Eastern (OR=0.31), Black (OR=0.32), Latin American (OR=0.28), Other Race/Ethnicity (a group those that identified as Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere, OR=0.22), and inviduals living in the West Health Region (OR=1.55).

Moreover, statistically-significant higher odd ratios were determined in the case of income and race; specifically, for individuals with a household income below CAD 25,000 that identified as Arab/Middle Eastern (OR=3.05), Black (3.19), Latin American (OR=2.80), and Other (4.59); and for individuals with a household income between CAD 25,000 and 59,999 that identified as Other racial minority (OR=6.93).

For Health Region and race, significant lower vaccination odds were identified for Black individuals in the Central Health Region (OR=0.44), and Other racial minorities and South Asian individuals that live in the West Health Region (ORs=0.41).

\footnotesize

\renewcommand{\arraystretch}{0.5}
```{r,model-table}
#| message: false
#| warning: false
#| echo: false
#| label: tbl-model
#| tbl-cap: "Multiple Regression Analysis-Predictors of Vaccination Status"

#logistic regression
source(here::here("code","regression.R"))

# create table with the model output
m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()%>%
  as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down" argument here.
  kableExtra::kable_styling(latex_options = "repeat_header")
```
