---
title: "Raking and model"
format: html
editor: visual
---

This document pertains the corrections on the clean dataset created in `Data_cleaning_and_covariates_Jan_05.qmd`. There are differences in the proportions of the dataset with regard to income, age groups, race/ethnicity, and geographical location when compared to the Census data. The dataset will be corrected using *raking* with the package `survey`.

# Raking

This section follows the steps provided by <https://sdaza.com/blog/2012/raking/> https://www.r-bloggers.com/2018/12/survey-raking-an-illustration/

The variables to define raking weights are:

-   income
-   age groups
-   geographical location
-   race and ethnicity

The population proportions in each case have been already presented in the data cleaning document.

The dataset has missing observations in the race (28% missing), and second_dose (26.3%) covariates. First attempt for raking and model will after removing all missing observations.

```{r,define-target-list}


library(weights)
library(here)
library(survey)
library(tidyverse)

clean_data<-read.csv(here("data","clean_dataset.csv"))

test<-clean_data%>%drop_na(race)

#remove municipal status as well to avoid problems

test<-test%>%
  select(-Municipal.status)

#make factors
test<-test %<>%
  mutate_at(c("income_ord","race","age_group","location","Geographic.area"),factor)

test<-test %>%
  mutate(first_dose_m=case_when(
    first_dose=="yes"~1,
    first_dose=="no"~0
  ))


#get percentages
wpct(test$income_ord)
wpct(test$age_group)
wpct(test$race)
wpct(test$location)

#create dataframes for weights

# t1<-svydesign(id=~1,data=test)
# 
# ages<-data.frame(age_group=c("16_24",
#                              "25_34",
#                              "35_44",
#                              "45_54",
#                              "55_64",
#                              "65_and_over"),
#                  Freq=c(0.127,
#                         0.129,
#                         0.128,
#                         0.149,
#                         0.137,
#                         0.167))
# 
# 
# 
# races<-data.frame(race=c("arab_middle_eastern",
#                          "black",
#                          "east_asian_pacific_islander",
#                          "indigenous",
#                          "latin_american",
#                          "mixed",
#                          "other",
#                          "south_asian",
#                          "white_caucasian"),
#                   Freq=c(0.016,
#                          0.047,
#                          0.066,
#                          0.028,
#                          0.015,
#                          0.01,
#                          0.052,
#                          0.087,
#                          0.678
#                   ))
# 
# income_ords<-data.frame(income_ord=c("under_15000",
#                                      "15000_24999",
#                                      "25000_39999",
#                                      "40000_59999",
#                                      "60000_89999",
#                                      "over_90000"),
#                         Freq=c(0.057,
#                                0.075,
#                                0.116,
#                                0.154,
#                                0.195,
#                                0.403
#                         ))
# 
# locations<-data.frame(location=c("All other cities",
#                                  "Toronto"),
#                       Freq=c(0.687,
#                              0.313))

########### use population totals ################
########## this is the approach used in the survey package examples #######


ages<-data.frame(age_group=c("16_24",
                             "25_34",
                             "35_44",
                             "45_54",
                             "55_64",
                             "65_and_over"),
                 Freq=c(1707959,
                        1734856,
                        1721407,
                        2003826,
                        1842444,
                        2245898))



races<-data.frame(race=c("arab_middle_eastern",
                         "black",
                         "east_asian_pacific_islander",
                         "indigenous",
                         "latin_american",
                         "mixed",
                         "other",
                         "south_asian",
                         "white_caucasian"),
                  Freq=c(212782,
                         638346,
                         886592,
                         376558,
                         197020,
                         130033,
                         705333,
                         1166361,
                         9118079 ))

income_ords<-data.frame(income_ord=c("under_15000",
                                     "15000_24999",
                                     "25000_39999",
                                     "40000_59999",
                                     "60000_89999",
                                     "over_90000"),
                        Freq=c(294643,
                               387688,
                               599624,
                               796052,
                               1007988,
                               2083176

                                  ))

locations<-data.frame(location=c("All other cities",
                                 "Toronto"),
                      Freq=c(10812589,
                              2635905
                                    ))



t.rake<-rake(t1,
             sample=list(~age_group,
                         ~race,
                         ~income_ord,
                         ~location),
             population=list(ages,races,income_ords,locations))


model<-svyglm(first_dose_m~income_ord+age_group+race,design=t1,family = quasibinomial(), control= list(maxit=50))

model.rake<-svyglm(first_dose_m~+Geographic.area*race,design=t.rake,family = quasibinomial(), control= list(maxit=50))

summary(model.rake)

model.test<-glmer(first_dose_m~Geographic.area*race+income_ord+age_group+Geographic.area,data=model.frame(t.rake),family=binomial)

```

Test example for raking from `survey` package

```{r,test-raking}
# data(api)
# 
# dclus1 <- svydesign(id=~dnum, weights=~pw, data=apiclus1, fpc=~fpc) #no clusters
# rclus1 <- as.svrepdesign(dclus1) #create replicate weights survey design
# 
# svymean(~api00, rclus1)
# svytotal(~enroll, rclus1)
# 
# pop.types <- data.frame(stype=c("E","H","M"), Freq=c(4421,755,1018))
# pop.schwide <- data.frame(sch.wide=c("No","Yes"), Freq=c(1072,5122))
# 
# rclus1r <- rake(rclus1, list(~stype,~sch.wide), list(pop.types, pop.schwide))
# 
# svymean(~api00, rclus1r)
# svytotal(~enroll, rclus1r)
test$first_dose<-as.factor(test$first_dose)

 ggplot(clean_data %>% count(first_dose,race) %>%    # G count number in each group
         mutate(pct=n/sum(n)),  # Calculate label positions
       aes(first_dose, n,fill=race)) +
  geom_bar(stat="identity") +
   facet_wrap(~as.factor(clean_data$Geographic.area))+
  theme_classic()
```
